---
title: "Explore PAHO Data: Dengue"
author: "Eugene Joh, MPH"
date: "April 8, 2018"
output: 
   html_document:
      toc: true #table of contents
      toc_float:
         collapsed: false
         smooth_scroll: true
      toc_depth: 3 #upto three depths of headings (specified by #, ## and ###)
      number_sections: false #if you want number sections at each table header
      theme: lumen #theme choices
      highlight: tango #specifies the syntax highlighting style
      code_folding: show #code folding
      df_print: paged #default print for dataframes in chunks
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

list.of.packages <- c("sf","sp","rgdal","maptools","rgeos","raster","ggmap","leaflet","ggmap", #mapping
                      "readr","readxl", #data import
                      "lubridate","tidyr","plyr","dplyr","data.table", "forcats", #data wrangling
                      "ggplot2","scales","grid","gridExtra", #data viz
                      "viridis","RColorBrewer","ochRe", #colour palette
                      "xts","forecast","zoo", #time series
                      "R.utils","wrapr","here","devtools" #utility
                      )

new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) { install.packages(new.packages) }
invisible(sapply(list.of.packages, require, character.only = T))
rm(new.packages,list.of.packages)
```

# Introduction
What is PAHO? Which regions do they cover? What is their mandate? What are they responsible for?

What is dengue? What is the vector? How is it transmitted? Who does it affect? What is the burden of dengue?

Why does PAHO have data on dengue? What are reportable diseases? Why is dengue a reportable disease? 

Terminology and Definitions:
What is a case? What is a fatality? What is incidence/prevalence/mortality? What is an epidemological week?

# Data 
What do these data files contain?
What relevant information do we want to look at?

## Import
downloaded flat-files named with tailing 'INIT', all deconstructed 
```{r data_import, echo=TRUE}
# Data Import ####
list.files(here('data'),pattern = "\\.csv$") #show all .csv files in /data folder
list.files(here('data'), pattern = 'INIT\\.csv$') #show all downloaded files in /data folder

# data import issues due to multibyte .csv encoding (UTF-16), the downloaded .csv were manually converted to singlebyte format
list.files(here('data'), pattern = 'FINAL\\.csv$') #show all modified files (due to multi-byte data import issues) in /data folder

# import 'FINAL' files into list ###
d.list <- lapply(seq_along(list.files(here('data'),pattern = 'FINAL\\.csv$')),
                    function(i) { #use readr read_csv()
                      readr::read_csv(here('data',list.files(here('data'),pattern = 'FINAL\\.csv$')[i]))
                    })
# add names to list import
names(d.list) <- list.files(here('data'), pattern = 'FINAL\\.csv$')
```

```{r epiweek_merge}
# merge the Epi Week data files (1-51 and 52) together
glimpse(d.list$`Epi_week_1-51_FINAL.csv`)
glimpse(d.list$Epi_week_52_FINAL.csv)
# check column names first
identical(names(d.list[[3]]),names(d.list[[4]])) #column names are the same

# merge the data together by usingbase::rbind()
d_epiwk <- rbind(d.list[[3]],d.list[[4]])
```

```{r mort_fatality}
# assign names for mortality and case fatality
d_mort <- d.list[[1]] #mortality data
d_fat <- d.list[[2]] #case fatality data
```

## Processing
conversion of object classes, any cleaning/processing steps
```{r change_names}
# change names where all white space is replaced with underscore
names(d_epiwk) <- gsub("\\s","_",names(d_epiwk))
```

```{r select_var}
# select relevant columns/variables
epiwk <- d_epiwk %>% #select columns of interest
  select(EW,Year, #dates
         Country,PAHO_Code,PAHO_Code_Num,ISO3166Alpha2,ISO3166Alpha3,Iso3166,Region, #locations
         Deaths,Lab_Confirm,Severe_Dengue,Suspected,Serotype,Type) %>% #case info
  arrange(Year,EW) #arrange by Year and EW
```

# Exploration
reporting descriptives (counts, proportions)

```{r na_vals}
# using lapply
lapply(epiwk, function(x) sum(is.na(x))) %>% 
  unlist() %>% data.frame(missing_values = .) %>% 
  mutate(name = rownames(.)) %>% 
  mutate(percent_missing = round(missing_values/nrow(.),1)) %>%
  select(name,missing_values,percent_missing)

# using purr::map()
purrr::map_df(epiwk,~sum(is.na(.))) %>% gather(key = name,value = missing_values) %>%
  mutate(percent_missing = round(missing_values/nrow(.),1))

```


# Data Visualization

```{r session_info}
sessionInfo()
```
