---
title: 'Explore PAHO Data: Dengue'
author: "Eugene Joh"
date: "`r format(Sys.Date(),'%B %d, %Y')`"
output: 
   html_document:
      theme: lumen
      highlight: tango
      code_folding: hide
      toc: true
      toc_float:
         collapsed: false
         smooth_scroll: true
      toc_depth: 3
      number_sections: false
params:
  i_seq: NULL
  i_names: NULL
  i_data: NULL
  
---

```{r setup, include=FALSE}
if (!require(here)) install.packages("here")
source(here::here('01_load_packages.R'))

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 9.2, fig.height = 7)

invisible(lapply(list.files(here::here("_fnc"),full.names = TRUE), source))
```

<!-- REMOVE THIS CHUNK AFTER PARAMETERIZATION -->
```{r test_param, eval = TRUE}
if (!require(here)) install.packages("here")

source(here::here('01_load_packages.R'))

source(here::here('02_data_import.R'))
```

## Exploratory Data Analysis
Exploratory data analysis (EDA) is a fundamental (but not glorified) process within in the "data science" pipeline. By conducting this step in a cognizant manner can help in understanding the structure of the data, plan effective data cleaning, and processing for the analysis portion of the workflow. A good starting point on EDA is this [R for Data Science chapter](http://r4ds.had.co.nz/exploratory-data-analysis.html). To aid the process of EDA, this **parameterized R markdown** file accepts a (relatively) "tidy" dataset and will return some high level descriptives that are core to understanding the dataset.

   - Data Type of each Column
   - Number of Missing Observations `NA`
   - Number of Unique Values

This can be easily be done by running the `glimpse` function from the `dplyr` package, but this markdown will provide some visualization to aid the understanding of the data. This will become useful when larger datasets (millions of rows and/or hundreds of columns) need to undergo some type of EDA.  

After identifying the structure of the data

## Data Types
```{r data_types}
check_data_types <- function(input, name) { # input is params$i_data
  
  df_dat_type <- map_df(input, typeof) %>% 
  gather(key = "Variable", value = "Data_Type") %>% 
  left_join(
    map_df(input, ~sum(is.na(.))) %>% 
      gather(key = "Variable", value = "N_Missing"),
    by = "Variable") %>% 
  left_join(
    map_df(input, ~length(unique(.))) %>% 
      gather(key = "Variable", value = "N_Unique"),
    by = "Variable") %>% 
  mutate(
      N_row = nrow(input),
      Data_File = name,
      Variable = fct_inorder(Variable)
    )
}

df_dat_type <- check_data_types(params$i_data, params$i_names)
df_dat_type %>% add_kable()
```

```{r p_data_types}
df_dat_type %>% 
  ggplot(aes(x = factor(fct_rev(fct_inorder(Variable))), y = N_Missing, fill = Data_Type, colour = Data_Type)) +
  geom_bar(stat='identity') + 
  labs(x = "Variable", y = "Counts of Missing",
       caption = paste0("Total Rows: ", df_dat_type$N_row[1])) +
  coord_flip() +
  facet_wrap(~Data_File, scales = "free")
```

## Summary Statistics

### Continuous
<!-- https://stackoverflow.com/questions/34594641/dplyr-summary-table-for-multiple-variables -->
```{r cont_summ}
tidy_summary(params$i_data) %>% add_kable()
```


### Categorical
```{r cat_list}
cat_list <- params$i_data %>% select_if(is.character) %>% 
  map(~count(data.frame(variable=.x), variable))
```

```{r program_chunks, results = 'asis', echo = TRUE}
for (i in seq_along(cat_list)) {
  
  cat('#', names(cat_list)[i], '\n')
  cat('```{r}', '\n')
  
  # code here
  eval(cat_list[i])
  
  cat('\n')
  
  cat('```', '\n')
  
}


```


```{r}
data("iris","airquality","mtcars")
my_list <- list(iris,airquality,mtcars)
```

```{r}
for (i in seq_along(my_list)) {
  
  cat('#', names(cat_list)[i], '\n')
  cat('```{r}', '\n')
  
  cat(')
  
  cat('\n')
  
  cat('```', '\n')
  
}
```



## Session Info
```{r sessinfo, echo=FALSE}
sessionInfo()
```

